---
title: "Analysis of Categorical Data"
author: "Ellen Chancey"
date: "July 23, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## One Sample Test  

```{r}
d <- data.frame(list(drink = c("Yes","No"), count = c(630, 670)))
str(d)
```
  
#### One  
Appropriate tests for categorical variables include chi-square, Fisher, and logistic regression.  
  
    
#### Two
```{r}
countsToCases <- function(x, countcol = "Freq") 
  { 
  idx <- rep.int(seq_len(nrow(x)), x[[countcol]]) 
  x[[countcol]] <- NULL 
  x[idx, ]
}

d.case <- countsToCases(d, countcol = "count")
str(d.case)
```
  
    
#### Three
```{r}
require(ggplot2)
qplot(d.case,main="Binge Drinking on Campus: A Survey of N = 1300 Undergraduates")
```
  
    
#### Four
```{r}
require(scales)
d.df <- data.frame(d.case)

ggplot(d.df, aes(x = d.case)) +
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels=percent, name="Percent") +
  scale_x_discrete(name="Response")
```
  
    
### Chi Square Goodness of Fit Test  
  
Five
```{r}
chisq.test(x=d$count, p=c(0.247,1-0.247))
```

```{r}
0.247*1300
(1-0.247)*1300
```
  
    
#### Six
```{r}
prop.test(x=630,n=1300,p=0.247,alternative="g")
```
There is not sufficient evidence to conclude that the rate of binge drinking is higher than 0.247 given that p = <2.2e-16 but the 95% confidence internval is 0.461:1.000

### Binomial Test  
  
#### Seven and Eight
```{r}
binom.test(x=630,n=1300,p=0.247,alternative="g")
```
There is not sufficient evidence to conclude that the rate of binge drinking is higher than 0.247 given that p = <2.2e-16 but the 95% confidence internval is 0.461:1.000

## Multi-Sample Tests  
  
#### One
```{r}
d <- read.csv("sales.csv",header=TRUE)
str(d)
library(MASS)
d.table <- xtabs(~Region + Sport,data = d)
summary(d.table)
```


#### Two
```{r}
prop.table(d.table,1)
mosaicplot(d.table)
```
  
  
#### Three  
Approprite test for multiple categorical variables include 2 sample t test, Mann-Whiteny, and Signed Rank  

### Chi-square Test of Independence  

#### Four and Five
```{r}
chisq.test(d.table)
```
There is sufficient evidence to suggest that the sales of sports gear at the retail outlet is dependent on geographic region (p = 9.473e-07).  

###Fisher's Exact Test  

#### Six and Seven
```{r}
fisher.test(d.table,workspace = 2e8)
```
There is sufficient evidence to suggest that the sales of sports gear at the retail outlet is dependent on geographic region (p = 3.106e-7).

## Dependent Sample Tests

```{r}
d <- matrix(c(13, 4, 25, 92),
              nrow = 2,
              dimnames = list("Case"=c("Exposed","Unexposed"),
                              "Control"=c("Exposed","Unexposed")))
```


### McNemar's Test

```{r}
require(exact2x2)
mcnemar.exact(d)
```
There is sufficient evidence to conclude that there is an association between exposure to the toxin and occurence of disease (odds ratio = 6.25, p value = 1.037e-4)