---
title: "Descriptive Statistics Homework"
author: "Ellen Chancey"
date: "July 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("babynames", lib.loc="~/R/win-library/3.4")
library("ggplot2", lib.loc="~/R/win-library/3.4")
library("Hmisc", lib.loc="~/R/win-library/3.4")
library("dplyr", lib.loc="~/R/win-library/3.4")
library("MASS", lib.loc="~/R/win-library/3.4")
library("vcd", lib.loc="~/R/win-library/3.4")
```

# Descriptive Statistics

### Describing a Dataset

One
```{r}
?survey()
```

Two
```{r}
str(survey)
```

Three
```{r}
describe(survey)
```

Four
```{r}
table(survey$Sex)
table(survey$Exer)
table(survey$Smoke)

table(survey$Sex, survey$Exer)
table(survey$Sex, survey$Smoke)

mytable<-table(survey$Sex, survey$Exer)
prop.table(mytable)

mytable2<-table(survey$Sex, survey$Smoke)
prop.table(mytable2)
```

Five
```{r}
# library(ggplot2)
ggplot(survey, aes(survey$Height,survey$Wr.Hnd, size=Age, color=Pulse)) +
    geom_point()
```

Six
```{r}
summary(survey) # contains mean
sd(survey$Wr.Hnd, na.rm=TRUE)
sd(survey$NW.Hnd, na.rm=TRUE)
sd(survey$Pulse, na.rm=TRUE)
sd(survey$Height, na.rm=TRUE)
sd(survey$Age, na.rm=TRUE)
```

Writing hand
```{r}
ggplot(survey, aes(x = survey$Sex, y = survey$Wr.Hnd )) + geom_boxplot()
```       

Non writing hand
```{r}
# ggplot(survey, aes(x = survey$Sex, y = survey$Nw.Hnd )) + geom_boxplot()
```

Pulse
```{r}
ggplot(survey, aes(x = survey$Sex, y = survey$Pulse )) + geom_boxplot()
```

Height
```{r}
ggplot(survey, aes(x = survey$Sex, y = survey$Height )) + geom_boxplot()
```

Age
```{r}
ggplot(survey, aes(x = survey$Sex, y = survey$Age )) + geom_boxplot()
```





#### Describing Bivariate Relationshops

One  
More males regularly or occasionally smoke, while more females report never smoking
```{r}
smoke2<-factor(survey$Smoke,levels=c("Never","Occas","Regul","Heavy"))

table(survey$Sex,smoke2)
```


```{r}
table(survey$Sex,smoke2)
smokebehave<-table(survey$Sex, smoke2)
prop.table(smokebehave)
```

Two  
The graph indicates that most students, whether male or female never smoke. It also indicates that there are fewer men who never smoke
```{r}
ggplot(data=survey,aes(x=smoke2)) +
  geom_bar() +
  facet_grid(.~ survey$Sex) +
  ggtitle("Smoking Behavior by Sex")
```

Three  
No, it appears the distribution of exercise for those who do smoke is immilar to those who don't.
```{r}
exer2<-factor(survey$Exer,levels=c("None","Some","Freq"))
smokeexer<-table(exer2,smoke2)
prop.table(smokeexer)
```

```{r}
ggplot(data=survey,aes(x=exer2))+
    geom_bar() +
    facet_grid(.~ survey$Smoke)
```

Four  
Most students have never smoked and do exercise
```{r}
mosaic(smokeexer)
```

Five
```{r}
ggplot(data=survey, aes(x=survey$Wr.Hnd, y=survey$Height)) +
  geom_point() +
  geom_smooth(method="lm")
```

Six  
This does not work
```{r}
cor(survey$Wr.Hnd, survey$Height, method = "pearson")
```

This works because it accounts for the missing values. The result indicates that there is a modest correlation
```{r}
cor(survey$Wr.Hnd, survey$Height, use="complete")

```

Seven  
They are consistent
```{r}
ff <- lm(survey$Height~survey$Wr.Hnd)
summary(ff)
```

Eight  
Estimate for Wr Hand is 3.1166

Nine  
Pulses of those who exercise frequently are generally a bit lower than those who exercise some or not at all
```{r}
ggplot(data=survey,aes(x=exer2, y = survey$Pulse)) +
  geom_boxplot()
```

Ten  
Data indicates that those who exervice frequently have a lower average pulse and that they also have a narrower standard deviation.
```{r}
?tapply()
tapply(survey$Pulse, exer2, mean, na.rm=TRUE)
tapply(survey$Pulse, exer2, sd, na.rm=TRUE)
```

Eleven  
The p value is 2.2e-16, so very small, based on this I would conclude that we can reject the hypothesis that there is no differene in pulse by exercice group
```{r}
ff <- aov(survey$Pulse~survey$Exer)
summary(ff)
```

Twelve  
Only Some and Frequent are significantly different because its p value is less than 0.05
```{r}
TukeyHSD(ff)
```


#### Annotations

```{r}
ggplot(economics, aes(date, unemploy)) + geom_line()
```

```{r}
ggplot(economics) +
  geom_rect(
   aes(xmin = start, xmax = end, fill = party),
   ymin = -Inf, ymax = Inf, alpha = 0.2, data = presidential
   ) +
  geom_vline(
   aes(xintercept = as.numeric(start)),
   data = presidential,
   colour = "grey50", alpha = 0.5
   ) +
  geom_text(
    aes(x = start, y = 2500, label = name),
    data = presidential,
    size = 3, vjust = 0, hjust = 0, nudge_x = 50
    ) +
  geom_line(aes(date, unemploy)) +
  scale_fill_manual(values = c("blue", "red"))
```

#### Additional Exercises

One  
Group = aes statement is required to use cyl, a continuous variable
```{r}
ggplot(data=mpg,aes(x=cyl, y=hwy, group= cyl)) +
  geom_boxplot()
```

Two  
Three bars
```{r}
ggplot(mpg, aes(drv)) + geom_bar(colour = "white")
```

Three bars with a number of sections, one for each hwy value
```{r}
ggplot(mpg, aes(drv, fill = hwy, group = hwy)) + geom_bar(colour = "white")
```

Three  bars with a large number of sections, one for each id
```{r}
mpg2 <- mpg %>% arrange(hwy) %>% mutate(id = seq_along(hwy))
ggplot(mpg2, aes(drv, fill = hwy, group = id)) + geom_bar(colour = "white")
```

Three  
Popularity of Jo has a relatively normal distributuion centered around the late 50s
```{r}
Jo <- dplyr::filter(babynames, name == "Jo")
ggplot(Jo, aes(year, n)) + geom_line()
```

my name's popularity has three waves of popularity, first around 1920 where it reaches 3000, then again in the 50s and 60s it has its highest peak up to 6000, then has a resurgence in the late 80s and early 90s (when I was born) around 1500
```{r}
Ellen <- dplyr::filter(babynames, name == "Ellen")
ggplot(Ellen, aes(year, n)) + geom_line()
```


## Document Information.

All of the statistical analyses in this document will be performed using `r R.version.string`.  R packages used will be maintained using the [packrat](http://rstudio.github.io/packrat/) dependency management system.  

```{r session-info}
sessionInfo()
```


