---
title: "Graphics in R"
author: "Ellen Chancey"
date: "July 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic Graphics in R

### Item 1
```{r}
?cars
dim(cars)
```

### Item 2
```{r}
plot(cars$speed,cars$dist)
```

### Item 3
```{r}
plot(cars$speed,cars$dist, 
    main="Scatterplot of Speed versus Distance",
    sub="Using plot() in R",
    xlab="Speed (miles per hour)",
    ylab="Stopping Distance (feet)",
    pch=19, # closed circle
    col="blue",
    col.axis="orange", # axis color
    col.lab="dark green", # axis label color
    bg="gray", # background color
    col.main="red", # main title color
    cex=1.5
    )
```

### Item 4
```{r}
plot(cars$speed,cars$dist, 
     main="Scatterplot of Speed versus Distance",
     sub="Using plot() in R",
     xlab="Speed (miles per hour)",
     ylab="Stopping Distance (feet)",
     pch=19, # closed circle
     col="blue",
     col.axis="orange", # axis color
     col.lab="dark green", # axis label color
     bg="gray", # background color
     col.main="red", # main title color
     cex=1.5, # sizing
     axes=FALSE # turns off auto axis
    )
abline(lm(dist~speed,data=cars)) # best fit line
grid() # adds grid
box(col="red",lwd=3,lty=3) # adds box around plot
legend("bottomright",  # adds a legend
       inset=0.01,
       title="Distance vs.Speed",
       legend=c("Observation"),
       col=c("blue"),
       pch=19,
       horiz=TRUE
       )
```


### Item 5
```{r}
library(ggplot2)
summary(diamonds)
par(mar = c(5, 7, 5, 5))
plot(diamonds$carat,
     diamonds$price,
     col=diamonds$color,
     pch=18,
     main = "Scatterplot of Diamond Size versus Price",
     xlab = "",
     ylab = "",
     xlim = c(0,5),
     ylim = c(0,20000),
     axes = FALSE
     )

axis(1, c(seq(from = 0, to = 5, by = 0.5)))
axis(2, c(seq(from = 0, to = 20000, by = 5000)), las = 1)
mtext("Size (carats)", side = 1, line = 3)
mtext("Price ($)", side = 2, line = 5)
legend("bottomright",
       title = "Diamond Color",
       legend = levels(diamonds$color),
       col = c(1:7),
       pch = 18,
       bty = "n"
       )
```

### Item 6
```{r}
qplot( diamonds$carat, 
       diamonds$price, 
       colour = diamonds$color,
       alpha = I(1/10),
       main = "Scatterplot of Diamond Sizes versus Price", 
       xlab = "Size (carats)",
       ylab = "Price ($)"
       )
```

###  Item 7
For diamonds under 1 carat, as size increases, price increases, For diamonds over 1 carat, they are more costly than smaller diamonds, but their price widely varies. Some diamond colors are more valuable than others, D is relatively inexpensive, E and F are moderately pricey and colors H, I, and J tend to be most expensive. 


# Advanced Graphics in R with ggplot

###  Item 1
Describe the data, aesthetic mappings and layers used.
The data contains fuel economy for 38 modles in 1998 and 2008.  
Uses a scatterplit to view the correlation between city and highway fuel efficiency (mpg)
```{r}
ggplot(mpg,aes(cty,hwy))+geom_point()

```

Diadmonds dataset contains price and size of 50,000 round cut diamonds. Graph uses a scatterplot to view relationship between carat and price.
```{r}
ggplot(diamonds,aes(carat,price))+geom_point()
?diamonds()
```

Economics dataset contains economic variables across time. This graphic looks at unemployment number over time in a line graph.
```{r}
ggplot(economics,aes(date,unemploy))+geom_line()
?economics
```

This histogram looks at the city fuel economy of models in the dataset.
```{r}
ggplot(mpg,aes(cty))+geom_histogram()
?mpg()
```

###  Item 2
Iris dataset contains measurements of 3 species of iris
```{r}
?iris
```

###  Item 3 
Map the data and aestetic mapping
```{r}
plot1 <- ggplot(data = iris, aes(Petal.Length, Petal.Width))
print(plot1)
```

###  Item 4 
Create the scatterplot
```{r}
plot2 <- plot1 + geom_point(aes(color = Species))
print(plot2)
```

###  Item 5 
Add linear regression
```{r}
plot3 <- plot2 + geom_smooth(method = "lm")
print(plot3)
```

###  Item 6 
Enhancements
```{r}
final <- plot3 +
  xlab("Petal Length (cm)") +
  ylab("Petal Width (cm)") +
  ggtitle("Petal Length versus Width")
plot(final)
```

###  Item 7 
Alter color
```{r}
plot1 <- ggplot(data = iris, aes(Petal.Length, Petal.Width))
print(plot1)

plot2 <- plot1 + geom_point(aes(color=Sepal.Length))
print(plot2)

# plot2 <- plot1 + geom_point(aes(color=Sepal.Length),aes(shape=Sepal.Width))
# print(plot2)

plot3 <- plot2 + geom_smooth(method = "lm")
print(plot3)

final <- plot3 +
  xlab("Petal Length (cm)") +
  ylab("Petal Width (cm)") +
  ggtitle("Petal Length versus Width")
plot(final)
```

A continuous variable cannot be mapped to shape, likely because there are too many items in a continuous variable a shape for each one can't be calculated because not all values can be calculated. The code below has been commented out so that it won't throw an error.
```{r}
# plot4 <- plot1 + geom_point(aes(shape=Sepal.Width))
# print(plot4)
```

This scatter plot uses sepal width as size
```{r}
plot5 <- plot1 + geom_point(aes(size=Sepal.Width))
print(plot5)
```

###  Item 8
```{r}
# install.packages(plyr)
```

###  Item 9
```{r}
library(plyr)
library(dplyr)
vert.line <- ddply(iris, .(Species), function(sp) {
  mean(sp$Petal.Length)})
```

###  Item 10
```{r}
ggplot(data = iris, aes(x = Petal.Length)) +
  geom_histogram() +
  geom_vline(data = vert.line, aes(xintercept = V1, color = Species)) +
  facet_grid(Species ~ . ) +
  ggtitle("Petal Lengths by Species")
```

###  Item 11
Moved Species to the other side of the dot within facet_grid()
```{r}
ggplot(data = iris, aes(x = Petal.Length)) +
  geom_histogram() +
  geom_vline(data = vert.line, aes(xintercept = V1, color = Species)) +
  facet_grid(. ~ Species) +
  ggtitle("Petal Lengths by Species")

```

###  Item 12
```{r}
ggplot(data = iris, aes(x=Petal.Length)) +
  geom_histogram(alpha = 0.2, position = "identity") +
  geom_vline(data = vert.line, aes(xintercept=V1 ,color=Species )) +
  ggtitle("Petal Length by Species Overlayed") +
  annotate("segment", x = 4.8, y = 14.5, xend = 4.8, yend = 5, size = 0.5,
             arrow = arrow(length=unit(.2, "cm"))) +
  annotate("text", label = "Overlap", x = 4.8, y = 15, size = 3,
             fontface = "bold")
```

###  Item 13
Using continuous variable for facet makes a large number of rows, making it very difficult to read.
```{r}
ggplot(data = iris, aes(x=Petal.Length)) +
  geom_histogram(alpha = 0.2, position = "identity") +
  geom_vline(data = vert.line, aes(xintercept=V1 ,color=Species )) +
  ggtitle("Petal Length by Species Overlayed") +
  annotate("segment", x = 4.8, y = 14.5, xend = 4.8, yend = 5, size = 0.5,
           arrow = arrow(length=unit(.2, "cm"))) +
  annotate("text", label = "Overlap", x = 4.8, y = 15, size = 3,
           fontface = "bold") +
  facet_grid(Petal.Width ~ .)
```

###  Item 14
```{r}
#?mpg()
#?aes()
```

a) drive train and fuel economy
```{r}
ggplot(data=mpg, aes(x=hwy,y=cty)) +
  ggtitle("The Relationship Between Fuel Economy and Drive Train") +
  #facet_grid(drv ~ .) +
  geom_point(aes(color=drv))
```

b) drive train, engine size, class
```{r}
ggplot(data=mpg, aes(x=class,y=displ)) +
  ggtitle("Relationship Between Drive Train, Class, and Engine Size") + 
  geom_point(aes(color=drv))
  #facet_grid(drv ~ .) 
```

c) fuel economy, engine size, cylinders  
Faceting by cyl doesn't significantly change the assessment of fuel economy and engine size because cylinders are closely grouped by engine size
```{r}
ggplot(data=mpg, aes(x=displ,y=cty)) +
  ggtitle("The Relationship Between Fuel Economy, Engine Size, and Cylinders") + 
  geom_point(aes(color=cyl)) +
  facet_grid( . ~ cyl)
```

###  Item 15
```{r}
ggplot(mpg, aes(displ, hwy)) + 
    geom_point() + 
    geom_smooth()
```

##### (a) 
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point() + geom_smooth(span = 0.2)
```

```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point() + geom_smooth(span = 1)
```

##### (b) 
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point() + geom_smooth(method="lm")
```

##### (c) 
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point() + geom_smooth(method="rlm")
```

### Item 16
```{r}
vignette("ggplot2-specs") 
```

```{r session-info}
sessionInfo()
```


