---
title: "R Programming Activity 2"
author: "Ellen Chancey"
date: "June 30, 2017"
output: 
  html_document:
    toc: true
    toc_float: true
  word_document: default
  pdf_document: default 
  odt_document: default
  md_document: default
  rtf_document: default
---

```{r setup, include=FALSE}
# This chunk of code sets up the R session to perform the analysis
# Load packages, load data, load any other source scripts that may contain
# code or objects you will want to run to produce the report

# load packages
library(knitr)
library(rmarkdown)
library(DT)

# declare global options
# knitr::opts_chunk$set(echo = FALSE)

# determine output format dynamically
out_type <- knitr::opts_knit$get("rmarkdown.pandoc.to")

# load data sets
# For example,

# define custom function for data label outputs
# The DT::datatable function is great for producing tables for HTML docs
# Otherwise, use the knitr::kable function to produce tables
# You should use the R help to learn about these two functions as they
# will need to be used to produce visually appealing tables for your 
# report

display_output <- function(dataset, out_type, filter_opt = 'none') {
  
  if (out_type == "html") {
    out_table <- DT::datatable(dataset, filter = filter_opt)
  } else {
    out_table <- knitr::kable(dataset)
  } 
  
  out_table
}
```

# Working with R Objects.

### Vector Manipulation.

Create a vector from a sequence of integers from 5 to 1:
```{r}
x <- c(5:1)
print(x)

```

Divide all elements of the vector x by 2:
```{r}
x/2

```

Square all elements of the vector x:
```{r}
x^2

```

Perform vector multiplication--multiply vector x by itself:
```{r}
x%*%x

```

Take the square root of the elements of the vector x:
```{r}
sqrt(x)

```

Find the minimum (smallest) element of the vector x:
```{r}
min(x)

```


### Matrix Manipulation.

Create a 2X2 matrix A and fill it (column-wise) with a sequence of integers from 1 to 4.
```{r}
A <- matrix(nrow=2,ncol=2,c(1:4))
print(A)

```

Create another 2x2 matrix B and a vector x
```{r}
B <- matrix(nrow=2,ncol=2,c(1:4)) 
print(B)
x <- c(2,2) 
```

Matrices can be added and subtracted.
```{r}
A+B
A - B 
```

A matrix can be transposed, more than once
```{r}
t(A) 
t(t(A)) 
```

Mathematical manipulation of matrices and  vectors
```{r}
A*B 
A%%B 
A%%x
x%*%A%*%x 
solve(A)
```

Assign aspects of a matrix, such as the row or column, to a vector
```{r}
C <- matrix(nrow=4,ncol=4,c(1:16)) 
C

a1 <- A[1,] 
a1

a2 <-A[2,]  
a2

b1 <- B[,1] 
b1

b2 <- B[,2] 
b2
```

Create a new matrix from an aspect of an existing matrix
```{r}
c_c <-C[,2:3]
c_c

c_r <-C[1:2,] 
c_r
```

Using repetition to population a matrix and returning only certain elements
```{r}
C[c(1,3),] <- matrix(nrow=2,ncol=4,c(rep(1,8)))
```

Replacing rows of a matrix
```{r}
C[c(1,3),]
```

Creating a matrix, returning it without the second row
```{r}
y <- matrix(c(4,5,2,3),nrow = 2) 
y[-2,] 
```

Adding amounts to columns of a matrix
```{r}
x<- matrix(nrow=3,ncol=2,c(1:6)) 
x

x+c(0,1)
```

Find the mean of each row, or column
```{r}
z <-matrix(nrow=3,ncol=2,c(1:6))
z

apply(z,1,mean)
apply(z,2,mean) 

```


### Character Strings.

Create a character object y that consists of the string "abc".
```{r}
y <- "abc"
```

Returns the length of the string y
```{r}
length(y)  
```

Returns the mode of storage for string y
```{r}
mode(y) 
```

Assigning multiple strings to a vector, finding length and mode of that vector
```{r}
z<- c("abc","29 88")
length(z) 
mode(z) 
```

Merging two strings
```{r}
w_w <-paste("Wonder", "Woman") 
```

Returning seperate strings
```{r}
ww <-strsplit(w_w," ") 
```

Evaluting the vector to determine if it is a list
```{r}
is.list(ww) 
```

Identifies which elements contain the specified string
```{r}
grep("West",c("North","South", "South East", "North West", "West")) 
```

Finding the number of characters in a string
```{r}
nchar("You're gonna need a bigger boat.") 
```

Finding specific characters within a string
```{r}
substr("Jaws",2,3) 
```

Finding the specific character where a substring begins in the string
```{r}
regexpr("aw","Jaws") 
```

Seperation of a string using a specified delimiter
```{r}
strsplit("6-16-2011", split="-")
```


### Sequences.

Returns examples of sequences
```{r}
example(seq)
```

Assigns the value 10 to n
```{r}
n <- 10 
```

Creates a sequence that starts with zero
```{r}
1:n-1 
```

Creates a sequence that starts with 1
```{r}
1:(n-1)
```

Creates a sequence that is however many numbers it takes to get rom the specified minimum to the specified maximum, using the specified interval
```{r}
seq(-10,10,by=2) -> sequence.1  
```

Creates a sequence with eleven elements, starting with the specified number and spacing by the specified amount
```{r}
sequence.2 <- seq(length=11, from = -10, by = 2) 
```

Creates a sequence that repeats the values in matrix x five times
```{r}
sequence.3 <- rep(x,5) 
```

Creates a sequence that repeats the listed numbers three times
```{r}
sequence.4 <- rep(c(5,2,7),3) 
```

Creates a sequence that repeats each listed number three times
```{r}
sequence.5 <- rep(c(5,2,7),each=3) 
```


### Logicals.

Creates an object, a, with four logical elements
```{r}
a <- c(TRUE,TRUE,FALSE,FALSE) 
```

Creates an object,b, with four logical elements
```{r}
b <- c(FALSE,TRUE,TRUE,FALSE) 
```

Returns a true if one exists in a or b at that position
```{r}
a|b 
```

Returns a true if both a & b are true at that position
```{r}
a&b 
```

Returns a false unless both a & b are true
```{r}
!a&b 
```

Returns a true if both a & b are false
```{r}
!a&!b 
```

Returns a false unless both a & b are true
```{r}
!(a&b)
```

Coerces a and b into numeric and adds the numbers
```{r}
a+b 
```

Evaluates matrix x elements and returns true if their value is larger than three
```{r}
dummy.1 <- x>3 
```

Evaluates matrix x elements and returns true if the value is equal to 3
```{r}
dummy.2 <- x ==3 
```

Reassigns x as a vector with elements one through ten
```{r}
x <- 1:10 
```

Evaluates if x contains any numbers larger than 8
```{r}
any(x>8) 
```

Evaluates if x contains any numbers larger than 88
```{r}
any(x>88) 
```

Evaluates if all elements in x are larger than 88
```{r}
all(x>88) 
```

Evaluates if all elements in x are larger than zero
```{r}
all(x>0) 
```

Reassigns x as a vector with elements one through three
```{r}
x <- 1:3 
```

Reassigns y as a vector with elements one, three, and four
```{r}
y<- c(1,3,4) 
```

Evaluates if the first element in x is equal to the first element in y and so on
```{r}
x==y 
```

Evaluates if all of the elements in x are equal by position
```{r}
all(x==y) 
```

Evaluates if there are difference between x and y
```{r}
identical(x,y) 
```

Reassigns x as a vector with elements one and two
```{r}
x <-1:2 
```

Reassigns y as a vector with elements one, three and four
```{r}
y<-c(1,3,4) 
```

Evaluates if elements in x and y are equal  
*Warns that x and y have differing lengths, but returns a false for the last element*
```{r}
x==y 
```

Evaluates if all the elements in x and y are equal  
*Returns false but warns about different lengths*
```{r}
all(x==y) 
```

Evaluates if and y are identical
```{r}
identical(x,y) 
```

Reassigns x to vector with elements one and two
```{r}
x<-1:2 
```

Reassigns y to vector with elements one and two
```{r}
y<-c(1,2) 
```

Returns the elements of x then the elements of y
```{r}
x;y 
```

Evaluates if x and y are identical, it is false
```{r}
identical(x,y) 
```

Returns type of storage for x, in thise case integer
```{r}
typeof(x) 
```

Returns type of storage for y, in this case double
```{r}
typeof(y) 
```


### Missing Data.

Creates a vector with elements one, two, three, and NA
```{r}
z <- c(1:3,NA) 
```

Function evaluates if the element in z is an NA and assigns that value to vector z.dummy
```{r}
z.dummy <-is.na(z) 
```

Does not evaluate if the elements in z are NA, creates an error message indicating is.na should be used
```{r}
z.dummy2 <- z==NA 
```

Returns NaN
```{r}
0/0 
```

Returns true 
```{r}
is.na(0/0) 
```

Returns true
```{r}
is.nan(0/0) 
```

Evaluates the elements of z for NaN  
*returns all false because NA is not NaN*
```{r}
is.nan(z) 
```

Returns NA because mean can't be determined
```{r}
mean(z) 
```

Indicates that the vector being evaluated contains an NA element, ignores it for calculating mean
```{r}
mean(z,na.rm=T) 
```

Reassigns x
```{r}
x <- c(88,NULL,12,168,13) 
```

Runs without error for the null element
```{r}
mean(x) 
```

Assigns u the element NULL
```{r}
u <- NULL 
```

Returns length of zero
```{r}
length(u) 
```

Assigns na to v
```{r}
v <- NA 
```

Returns length one
```{r}
length(v) 
```


### Simple Subsetting.

Creates a vector with elements one through five, then 3 elements of NA
```{r}
x <- c(c(1:5),c(rep(NA,3))) 
```

Evaluates if the elements in x are NA, the last three elements are true
```{r}
is.na(x) 
```

Returns elements of x that are evaluated as being na
```{r}
x[is.na(x)] 
```

Returns all elements of x that are not na
```{r}
x[!is.na(x)] 
```

Evaluates if each element is greater than two, returns NA for final three NA elements
```{r}
x>2 
```

A logical evaluation of the elements in x  based on if they are not na and are larger than two, returns NA for NA elements
```{r}
!is.na(x)&x>2 
```

A logical evaluation of the elements in x based on if they are not null and are larger than two, returns FALSE for NAs
```{r}
x[!is.na(x)&x>2] 
```

Reassigns x to a 2X3 matrix containing values one through six
```{r}
x<- matrix(1:6,2,3) 
```

Creates a list with componenets of different classes
```{r}
w <- list(one=1:5, two = 0.6, three = "HI") 
```

Returns the first component in the list, a vector with elements one through five
```{r}
w$one 
```

Returns the first component in the list, a vector with elements one through five
```{r}
w[["one"]] 
```

Returns the first component in the list, a vector with elements one through five
```{r}
w[[1]] 
```

Returns the third element of the first componenet in the list
```{r}
w[[c(1,3)]] 
```

Returns the third element of the first component
```{r}
w[[1]][[3]] 
```

Returns the first and third componenet
```{r}
w[c(1,3)] 
```

Reassigns z as a vector with four elements
```{r}
z<- c(5,2,-3,8) 
```

Returns the elements that make up the vector
```{r}
z 
```

Returns TRUE if the value of multiplying the element by itself results in a number larger than 8
```{r}
z*z > 8 
```

Returms elements of the vector z that meet the specified criteria
```{r}
z[z*z>8] 
```

Assigns the values that meet the specified criteria to the vector w
```{r}
w <-z[z*z>8] 
w
```

Reassigns x as a vector with six elements
```{r}
x <- c(6,1:3,NA,12) 
x
```

Returns the values in vector x that meet the criteria, **including** NA
```{r}
x[x>5] 
```

Returns the values in vector x that meet the criteria, **excluding** NA
```{r}
subset(x,x>5) 
```

Changes any element in x meeting the requirement to the value specified
```{r}
x[x>3] <- 0 
x
```

Returns the position number of the elements that meet the requirement
```{r}
which(z*z>8) 
```


### Arrays.
### Write the Code.

####1. Create a vector of length 10 and fill it with a sequence of integers.
```{r}
v <- c(12:21)
v
```

####2. Coerce the vector into a matrix of 5 rows and 2 columns.
```{r}
vm <-matrix(v,nrow=5,ncol=2)
vm
```

####3.  Name the columns of the matrix "A" and "B".
```{r}
colnames(vm) <- c("A","B")
vm
```

####4. Generate 12 random numbers from a uniform distribution and store them in a matrix of 4 rows and 3 columns.
```{r}
d <- rnorm(12,6)
d
dm <-matrix(d,nrow=4,ncol=3)
dm
```

####5. Name the rows of the matrix "1st Row", "2nd Row", "3rd Row", and "4th Row".
```{r}
rownames(dm) <-c("1st Row","2nd Row","3rd Row", "4th Row")
dm
```

####6. Using bracket notation, print the element of the matrix that is in the 2nd row, 1st column. Assign the entire 2nd column of the matrix to a new object r.
```{r}
dm[2,1]
r<-dm[,2]
r
```

####7. Assume that we have recorded the names and ages for four people
James is 27, Art is 42, Kate is 29, and Alex is 33. Create a vector of names and a vector of ages from the data, making sure that you keep the ordering of the elements consistent.
```{r}
n<-c("James", "Art","Kate","Alex")
n
a<-c(27,42,29,33)
a
```

Use the class() function to print the class metadata R has stored for each vector.
```{r}
class(n)
class(a)
```

Using data.frame(), combine the two vectors into a dataframe and name the columns something informative. Print out the dataframe.
```{r}
df<-data.frame(name=n,age=a)
df
```

####8. Assume that we have the following observations of temperature:23???C, 30???, 19???, 27???.
Create a vector C concatenating these values. 
```{r}
C<-c(23,30,19,27)
C
```

Recall the relationship between Celsius and Fahrenheit temperatures: ???F = ???C  ×5 + 32.
Using a single line of code that includes arithmetic operations, create a new vector F that contains the temperatures transformed to Fahrenheit scale.
```{r}
F<-(C*(9/5)+32)
F
```


####9. Suppose we have a sample of 15 plants from all across the country and their state of origin is specified by a character vector:
```{r}
state<- c("MO","KS","KS","KS","MO","NE","OK","MO","KS","NE","NE","OK","MO","KS","KS")
state
```

Use the factor() function to change the character vector state to a factor and produce a frequency table that represents the number of observations in our sample from each state. Use the levels option within factor() to reorder the state categories to: Oklahoma, Kansas, Missouri, Nebraska.
```{r}
table(state) # levels appear in alphabetical order

stlvl<-c("OK","KS","MO","NE") 
fstate<-factor(state,stlvl)
table(fstate) # order specified in level vector
```

####10. Suppose, further, that we have a measure of production for the plans in another vector:
```{r}
prod<-c(2341,9873,23933,1999,7214,9089,1099,2389,5435,5757,
        +2521,7333,21909,15110,4369)
prod
```

Use the help to investigate the tapply() function. 
```{r}
?tapply()
```

Write the command line that will produce the mean production for each state using tapply().
```{r}
tapply(prod,fstate,mean) # vector, level of detail, function to be applied
```

Write similar command lines that will produce the minimum and maximum production for each state.
```{r}
tapply(prod,fstate,min)
tapply(prod,fstate,max)
```

##### Document Information.
All of the statistical analyses in this document will be performed using `r R.version.string`.  R packages used will be maintained using the [packrat](http://rstudio.github.io/packrat/) dependency management system.  

```{r session-info}
sessionInfo()
```

