---
title: "Common Probability Distributions"
author: "Ellen Chancey"
date: "July 2, 2017"
output: 
  html_document:
    toc: true
    toc_float: true
  word_document: default
  pdf_document: default 
  odt_document: default
  md_document: default
  rtf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(rmarkdown)
library(graphics)
out_type <- knitr::opts_knit$get("rmarkdown.pandoc.to")
display_output <- function(dataset, out_type, filter_opt = 'none') {
  
  if (out_type == "html") {
    out_table <- DT::datatable(dataset, filter = filter_opt)
  } else {
    out_table <- knitr::kable(dataset)
  } 
  
  out_table
}
```
## Discrete Distributions

### Bernoulli
The **Bernoulli distribution**, named for [Jacob Bernoulli](https://en.wikipedia.org/wiki/Jacob_Bernoulli), assigns probability to the outcomes of a single Bernoulli experiment---one where the only possible outcomes can be thought of as a "success" or a "failure" (e.g., a coin toss). Here, the random variable $x$ can take on the values 1 (success) with probability $p$, or 0 (failure) with probability $q = 1 - p$.  The plot below contains the pmf of two Bernoulli distributions.  The first (in gray) has a probability of success $p = 0.2$ and the second (in black) has a probability of success $p = 0.5$.  

```{r}
x <- 0:1
plot(x, dbinom(x, 1, 0.2), type = "h", ylab = "f(x)", ylim = c(0,1), lwd = 8, col = "dark gray",
      main = "Bernoulli(0.2)")
lines(x, dbinom(x, 1, 0.5), type = "h", lwd = 2, col = "black")
legend(0.7, 1.0, c("Bernoulli(0.2)","Bernoulli(0.5)"),col=c("dark gray","black"), lwd=c(8,2))

```

The Bernoulli experiment forms the foundation for many of the next discrete distributions.


### Binomial
The **binomial distribution** applies when we perform $n$ Bernoulli experiments and are interested in the total number of "successes" observed.  The outcome here, $y = \sum x_i$, where $P(x_i = 1) = p$ and $P(x_i = 0) = 1 - p$.  The plot below displays three binomial distributions, all for $n = 10$ Bernoulli trials: in gray, $p = 0.5$; in blue, $p = 0.1$; and in green, $p = 0.9$.  
```{r}
x <- seq(0, 10, 1)
plot(x, dbinom(x, 10, 0.5), type = "h", ylab = "f(x)", lwd = 8, col = "dark gray", ylim = c(0,0.5),
     main = "Binomial(10, 0.5) pmf")
lines(x, dbinom(x, 10, 0.1), type = "h", lwd = 2, col = "blue")
lines(x, dbinom(x, 10, 0.9), type = "h", lwd = 2, col = "green")
legend(3, 0.5, c("Binomial(10,0.1)","Binomial(10,0.5)","Binomial(10,0.9)"), col=c("blue","dark gray","green"),lwd=c(2,8,2))

```

We can see the shifting of probability from low values for $p = 0.1$ to high values for $p = 0.9$.  This makes sense, as it becomes more likely with $p = 0.9$ to observe a success for an individual trial.  Thus, in 10 trials, more successes (e.g., 8, 9, or 10) are likely.  For $p = 0.5$, the number of successes are likely to be around 5 (e.g., half of the 10 trials).

### Hypergeometric
In the example I have below, I have set the number of balls in the urn to 10, 5 of which are white and 5 of which are black.  I have also fixed the number of balls drawn from the urn to 5.  

Play around with the parameters and describe what you see:  
When 5 of each color and 5 are drawn, highest probability of 2 and 3. 
```{r}
x <- seq(0, 10, 1)
plot(x, dhyper(x, 5, 5, 5), type = "h", ylab = "f(x)", lwd = 2,
    main = "Hypergeometric(5,10,5) pmf")

```

When the number of white balls increases, the distribution increases. 
```{r}
x <- seq(0, 10, 1)
plot(x, dhyper(x, 7, 3, 5), type = "h", ylab = "f(x)", lwd = 2,
    main = "Hypergeometric(5,10,5) pmf")

```

When the number of black balls increases, the distribution decreases.
```{r}
x <- seq(0, 10, 1)
plot(x, dhyper(x, 3, 7, 5), type = "h", ylab = "f(x)", lwd = 2,
    main = "Hypergeometric(5,10,5) pmf")

```

When the number of draws increases, the distribution narrows and increases.
```{r}
x <- seq(0, 10, 1)
plot(x, dhyper(x, 5, 5, 8), type = "h", ylab = "f(x)", lwd = 2,
    main = "Hypergeometric(5,10,5) pmf")

```
 
### Poisson
What happens if you increase $\lambda$? To 2? To 3?

```{r}
x <- seq(0, 5, 1)
plot(x, dpois(x, 1), type = "h", ylab = "f(x)", main = "Poisson(1) pmf", lwd = 2)

```

The distribution moves higher as $\lambda$ increases
```{r}
x <- seq(0, 5, 1)
plot(x, dpois(x, 2), type = "h", ylab = "f(x)", main = "Poisson(2) pmf", lwd = 2)

```

### Geometric
What happens to the geometric distrbution if you vary $p$?  Show me a few plots and explain.

When probability increases, it takes fewer numbers of trials to get a success so the distribution gets narrower.

######Oringinal with 0.2 distribution
```{r}
x <- seq(0, 20, 1)
plot(x, dgeom(x, 0.2), type = "h", ylab = "f(x)", lwd = 2,
    main = "Geometric(0.2) pmf")

```

######Revised with 0.5 distribution
```{r}
x <- seq(0, 20, 1)
plot(x, dgeom(x, 0.5), type = "h", ylab = "f(x)", lwd = 2,
    main = "Revised Geometric(0.5) pmf")

```

######Revised with 0.8 distribution
```{r}
x <- seq(0, 20, 1)
plot(x, dgeom(x, 0.8), type = "h", ylab = "f(x)", lwd = 2,
    main = "Revised Geometric(0.8) pmf")

```


### Negative Binomial
The negative binomial I have below has set $r = 1$, so it's identical to the geometric above.  Play around with $r$ and see how it changes.

```{r}
x <- seq(0, 20, 1)
plot(x, dnbinom(x,1,0.2), type = "h", ylab = "f(x)", lwd = 2,
     main = "Negative Binomial(0.2) pmf")

```

$r = 4$ and probability = 0.4
```{r}
x <- seq(0, 20, 1)
plot(x, dnbinom(x,4,0.4), type = "h", ylab = "f(x)", lwd = 2,
     main = "Negative Binomial(0.4) pmf")

```

## Continuous Distributions

### Exponential 
Vary $\lambda$ and describe.  
With a higher rate, the probability shifts down, making it more likely to occur early
```{r}
x <- seq(0, 10, 0.01)
plot(x, dexp(x, 1), type = "l", ylab = "f(x)", lwd = 2, 
    main = "Exponential(1) pdf")

```

```{r}
x <- seq(0, 10, 0.01)
plot(x, dexp(x, 4), type = "l", ylab = "f(x)", lwd = 2, 
    main = "Exponential(4) pdf")

```

### Normal
Vary $\sigma$ and see how the distribution changes.  If you make it too big, you may need to adjust the *x*-axis by making the sequence span a wider range than $-5$ to $5$.  You can use a trial-and-error approach to determing the proper limits for x for a given $\sigma$.

$\sigma$ of 1
```{r}
x <- seq(-5, 5, 0.01)
plot(x, dnorm(x, 0, 1), type = "l", ylab = "f(x)", main = "Normal(0, 1) pdf")

```

$\sigma$ of 0.5
```{r}
x <- seq(-5, 5, 0.01)
plot(x, dnorm(x, 0, .5), type = "l", ylab = "f(x)", main = "Normal(0, 0.5) pdf")

```

$\sigma$ of 3
```{r}
x <- seq(-5, 5, 0.01)
plot(x, dnorm(x, 0, 3), type = "l", ylab = "f(x)", main = "Normal(0, 3) pdf")

```

### Chisquare
How do the degrees of freedom change the shape? Plot a few and explain.  

Degrees of Freedom: 3  
When DOF are low, as in this plot, the highest probability occurs early. The probability peaks quickly and the reduces at a slightly lower rate.
```{r}
x <- seq(0, 20, 0.01)
plot(x, dchisq(x, 3), type = "l", ylab = "f(x)", main = "Chi-square(3) pdf")

```

Degrees of Freedom: 6  
When DOF are in a medium range, as in this plot, the highest probability occurs after a handful of observations so that probability steadily peaks and then gradually reduces after.
```{r}
x <- seq(0, 20, 0.01)
plot(x, dchisq(x, 8), type = "l", ylab = "f(x)", main = "Chi-square(8) pdf")

```

Degrees of Freedom: 15  
When DOF is large, as in this plot, probability builds slowly with a wide peak.
```{r}
x <- seq(0, 20, 0.01)
plot(x, dchisq(x, 15), type = "l", ylab = "f(x)", main = "Chi-square(15) pdf")

```

### Students t

How do the degrees of freedom change the shape?  Plot a few and explain.  

The DOF on the t test does not alter the distribution, as demonstrated in the following plots which have DOF of 2, 6, and 10.
```{r}
x <- seq(-5, 5, 0.01)
plot(x, dt(x, 2), type = "l", ylab = "f(x)", main = "Student's t(2) pdf")

```

```{r}
x <- seq(-5, 5, 0.01)
plot(x, dt(x, 6), type = "l", ylab = "f(x)", main = "Student's t(6) pdf")

```

```{r}
x <- seq(-5, 5, 0.01)
plot(x, dt(x, 10), type = "l", ylab = "f(x)", main = "Student's t(10) pdf")

```

### F

How do the degrees of freedom (numerator and/or denominator) change the shape?  Plot a few and explain.

```{r}
x <- seq(0, 6, 0.01)
plot(x, df(x, 12, 15), type = "l", ylab = "f(x)", main = "F(12, 15) pdf")

```

In this example, with dof1 reduced down to 3, but dof2 remaining at 15, the distribution gets wider and the rate of change early on increases more dramatically
```{r}
x <- seq(0, 6, 0.01)
plot(x, df(x, 3, 15), type = "l", ylab = "f(x)", main = "F(3, 15) pdf")

```

In this example, with dof2 increased to 25, there is not a lot of change from the original F(12, 15) plot. 
```{r}
x <- seq(0, 6, 0.01)
plot(x, df(x, 12, 25), type = "l", ylab = "f(x)", main = "F(12, 25) pdf")

```

Similarly in this example, with dof2 decreased to 5, there is not a lot of change from the original F(12, 15) plot. On this plot, you can see that the peak decreases more slowly than the F(12,25)
```{r}
x <- seq(0, 6, 0.01)
plot(x, df(x, 12, 5), type = "l", ylab = "f(x)", main = "F(12, 5) pdf")

```

## Document Information.

All of the statistical analyses in this document will be performed using `r R.version.string`.  R packages used will be maintained using the [packrat](http://rstudio.github.io/packrat/) dependency management system.  

```{r session-info}
sessionInfo()
```
